%% SEA CHAPTER 9 - OPTIMIZATION IN DESIGN AND OPERATIONS
% SEA Question Location in \label{sea-Chapter#-Problem#}
% ANSWERS NEED UPDATING
\begin{exercises}
    \begin{exercise}
    \label{sea-9-1}
        Specify the dimensions of the sides of a rectangle of perimeter $p$ so that the area it encloses will be maximum.
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-9-2}
        The cost per unit produced at a certain facility is represented by the function
        \begin{equation}
            UC = 2x^2-10x+50
        \end{equation}
        where $x$ is in thousands of units produced. For what value of $x$ would unit cost be minimized (other than zero)? What is the minimum cost at this volume? Show that the value found is truly a minimum.
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-9-3}
        Advertising expenditures have been found to relate to profit approximately in accordance with the function
        \begin{equation}
            P = x^3-100x^2+3,125x
        \end{equation}
        where $x$ is the expenditure in thousands of dollars. What advertising expenditure would produce the maximum profit? What profit is expected at this expenditure? Show that the derived result is truly a maximum.
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-9-4}
        The cost of producing and selling a certain item is for the first 1,000 units, for a production range between 1,000 and 2,500 units, and $\$205x-\$97.500$ for more than 2,500 units, where $x$ is the number of units produced. If the selling price is \$200 per unit, and all units produced are sold,find the level of production that will maximize profit.
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-9-5}
        Ethyl acetate is made from acetic acid and ethyl alcohol. Let $x=$ pounds of acetic acid input, $y=$ of ethyl alcohol input, and $z=$ of ethyl acetate output. The relationship of output to input is
        \begin{equation}
            \frac{z^2}{(1.47x-z)(1.91y-z)}=3.9
        \end{equation}
        \begin{enumerate}[label=\alph*)]
            \item Determine the output of ethyl acetate per pound of acetic acid,where the ratio of acetic acid of ethyl alcohol is 2.0,1.0,and 0.67,and graph the result.
            \item Graph the cost of material per pound of ethyl acetate for each of the ratios given and determine the ratio for which the material cost per pound of ethyl acetate is a minimum if acetic acid costs \$0.80 per pound and ethyl alcohol costs \$0.92 per pound.
        \end{enumerate}
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-9-6}
        It has been found that the heat loss through the ceiling of a building is 0.13 Btu per hour per square foot of area per degree Fahrenheit. If the $2,200ft^2$ ceiling is insulated,the heat loss in Btu per hour per degree temperature difference per square foot of area is taken as
        \begin{equation}
            \frac{1}{(\frac{1}{0.13})(\frac{t}{0.27})}
        \end{equation}
        where tis the thickness in inches.The in-place cost of insulation 2,4,and 6 in thick is \$0.18, \$0.30,and \$0.44 per square foot,respectively.The building is heated to 75$^{\circ}$F 3,000 hrs per year by a gas furnace with an efficiency of 50\%.The mean outside temperature is 45$^{\circ}$F and the natural gas used in the furnace costs \$4.40 per and has a heating value of 2,000 Btu per What thickness of insulation, if any, should be used if the interest rate is 10\% and the resale value of the building 6 years hence is enhanced \$850 if insulation is added,regardless of the thickness?
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-9-7}
        An overpass is being considered for a certain crossing. The superstructure design under consideration will be made of reinforced concrete and will have a weight per foot depending on the span between piers in accordance with $W=32(S)+1,850$. Piers will be made of steel and will cost \$250,000 each. The superstructure will be erected at a cost of \$3.20 per pound. If the number of piers required is to be one less than the number of spans, find the number of piers that will result in a minimum total cost for piers and superstructure if $L=1,250$ft. 
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-9-8}
        Two girder designs are under consideration for a bridge for a 1,200-foot crossing.The first is expected to result in a superstructure weight per foot of $22(S)+800$, where $S$ is the span between piers. The second should result in superstructure weight per foot of $20(S)+1,000$. Piers and two required abutments are estimated to cost \$220,000 each. The superstructure will be erected at a cost of \$0.55 per pound. Choose the girder design that will result in a minimum cost and specify the optimum number of piers.
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-9-9}
        What is the cost advantage of choosing the best girder design for the bridge described in \ref{sea-9-8}? If the number of piers is determined from the best girder design alternative, but the other design alternative is adopted, what cost penalty is incurred?
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-9-10}
        A used automobile can be purchased by a student to provide transportation to and from school for \$5,500 as is (i.e.,the auto will have no warranty). First-year maintenance cost is expected to be \$350 and the maintenance costs will increase by \$100 per year thereafter. Operation costs for the automobile will be \$1,200 for every year the auto is used and its salvage value decreases by 15\% per year.
        \begin{enumerate}[label=\alph*)]
            \item What is the economic life without considering the time value of money?
            \item With interest at 16\%,what is the economic life?
        \end{enumerate}
    \end{exercise}
    \begin{solution}
    \end{solution}
\end{exercises}
% SKIPPED