%% SEA CHAPTER 4 - PRELIMINARY SYSTEM DESIGN
% SEA Question Location in \label{sea-Chapter#-Problem#}
% NEEDS UPDATING
\begin{exercises}
    \begin{exercise}
    \label{sea-4-1}
        Select a system of your choice and develop operational functional flow block diagrams (FFBDs) to the third level. Select one of the functional blocks and develop maintenance functional flows to the second level.Show how the maintenance functional flow diagrams evolve from the operational flows.
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-4-2}
        Describe how specific resource requirements (i.e., hardware, software, people, facilities, data,and elements of support) are derived from the functional analysis.
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-4-3}
        Describe the steps involved in transitioning from the functional analysis to a “packaging scheme”for the system.Provide an example.
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-4-4}
        Refer to the allocation in Figure 4.6. Explain how the quantitative factors (i.e.,TPMs) at the unit level and assembly level were derived.
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-4-5}
        What steps would you take in accomplishing the allocation of requirements for a system-of-systems (SOS?) configuration? Describe the overall process.
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-4-6}
        Select a system of your choice and assign some top-level TPMs. Allocate these requirements as appropriate to the second and third levels.
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-4-7}
        Refer to Figure 4.7. How would you define the design-to requirements for the common unit?"
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-4-8}
        Describe what is meant by interoperability? Why is it important?
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-4-9}
        Describe what is meant by environmental sustainability? Identify some specific objectives in the design for such. 
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-4-10}
        Why is the design for security important? Identify some specific design objectives. 
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-4-11}
        Why is the design for security important? Identify some specific design objectives. 
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-4-12}
        Refer to Figure 4.1. Why is the development of a specification tree important? What inherent characteristics should be included?
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-4-13}
        Refer to Figure 4.2. How are metrics established for the function shown? Give an example.
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-4-14}
        Refer to Figure 4.4. Describe some of the interfaces or interactions that must occur to ensure a completely integrated approach in the development of the hardware, software, and human system requirements. Be specific.
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-4-15}
        Define CAD, CAM, CAS, Macro-CAD, and their interrelationships. Include an illustration showing interfaces, information/data flow, etc.
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-4-16}
        Assume that you are a design engineer and are looking for some analytical models/tools to aid you in the synthesis, analysis, and evaluation process. Develop the criteria that you would apply in selecting the most appropriate tools for your application.
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-4-17}
        Assume that you have selected an analytical model for a specific application. Explain how you would validate that the model is adequate for the application in question.
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-4-19}
        List some of the benefits that can be derived from the use of computer-based models. Identify some of the concerns associated with the application of such.
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-4-20}
        Refer to Figure 4.8. Identify some of the objectives in selecting the appropriate methods/tools for accomplishment of the tasks in the right-hand columns.
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-4-21}
        Refer to Figure 4.12. Identify some of the objectives in designing a “tool set” as shown by the seven blocks in the figure.
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-4-22}
        Briefly identify some inputs and outputs for the system design review, the equipment/software design review, and the critical design review.
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-4-23}
        On what basis are formal design reviews scheduled? What are some of the benefits derived from conducting formal design reviews?
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-4-24}
        How are supplier requirements determined? How are these requirements passed on to the supplier?
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-4-25}
        Design is a team effort. Explain why! How should the design team be structured? How does systems engineering fit into the process?
    \end{exercise}
    \begin{solution}
    \end{solution}
    
    \begin{exercise}
    \label{sea-4-26}
        What is the desired output of the Preliminary System Design Phase?
    \end{exercise}
    \begin{solution}
    \end{solution}
\end{exercises}
% SKIPPED