\section{Summary and Extensions}\index{Summary and Extensions}

There is not unanimity around the definition of Model-based systems engineering in the market place and most uses of the term Model-based systems engineering are not as broad as what is addressed here. The same iterative approach that is used to define and design a new system (the ``top-down'' engineering approach) can be used to analyze and improve an existing system (``middle-out'' or ``reverse'' engineering). By intentionally adopting a ``broad'' definition of a system the approach can be used across the widest possible spectrum of systems to be analyzed or constructed. Clearly, from the definitions and discussion, the aim of a system model is to be able to analyze and gain insight into real systems whether human-made or not. With human-made or ``engineered'' systems, the aim is to find a realizable solution to a stated need using effective engineering and management processes.

The chapters that follow will not go into exhaustive detail but will strive to provide a high-level view of the Model-based systems engineering process. Various aspects of the iterative systems engineering process will be addressed to develop a clear understanding of ``what is model-based systems engineering'' and ``what are its advantages?'' Because of its very inter-relatedness it is difficult to determine the order in which discuss MBSE’s concepts. The discussion here will begin with language, the foundational blocks of the systems engineering process. From there it will turn to the model, the process and the functional and physical topics. Along the way, most of the major concepts of MBSE will be introduced and explained. It is the hope of the creative team who write this Primer that it will prove a useful introduction to the world of problem solving with MBSE.

%%----------------------------------------------------------------------------------------
%	PROBLEMS
%%----------------------------------------------------------------------------------------

\input{Problems/Problems07}